How to be a good member of a scientific software community [Article v\versionnumber]

### 2021-07-13 paper.tex
more text

In all of these cases, it is necessary to interact with the people who develop
and support the code. How you go about it has an enormous impact on your
likelihood of success and how you are viewed, and asking complete strangers for
help is often intimidating, especially the first time you do it. However, the
best way to do so is rarely taught explicitly --- at best, it's something one
picks up by example, from fellow lab members or the PI.  {\emph The goal of this
paper is to reveal the hidden curriculum -- what's expected of a software
consumer, how to ask for help, how to contribute productively to a software
community (regardless of whether you can write code).}

This article is written from a perspective of academic open-source scientific
software. Much of what we suggest is universal, but some points --- like the
reminder that the people providing the support are likely volunteers --- are
not.  Regardless, we view the interaction between developers and users as an
informal social contract, where each has obligations and expectations for the
other. Obviously, every software community is unique, and many have specific
conventions for interaction, but we hope that the recommendations we make are at
the very least a good starting point for new users of scientific software.

@@

Finally, you can consider diving into the code. This decision depends very much
on the complexity of the package, your skill level, and how urgently you need
the solution. More often than not, comprehending the whole software package will
be too time consuming, but there are tricks and shortcuts one can use to at
least figure out where things might be going wrong. Figuring out where the error
is occurring can sometimes help you figure out why it happened, so examine the
stack trace if there is one. Alternatively, you can search the code base for the
error message, to see what drove the error; it's not uncommon for the comments
in the code to be clearer and more indicative of what could go wrong than the
error message itself. To be clear: as an end user, \emph{you are not obligated
to check the code}, but it can sometimes be a good shortcut to help you either
solve your problem on your own or give the developers what they need to solve
it.

@@

@@

\begin{checklist}{Good community member}
\begin{itemize}
\item Tries to solve problem themselves first
\item Asks for help in the right place
\item Writes informative bug reports
\item Cites and acknowledges software appropriately
\item Treats fellow members and developers with courtesy and respect
\item Contributes to the community
\end{itemize}
\end{checklist}

>>>>

In all of these cases, it is necessary to interact with the people who develop
and support the code. How you go about it has an enormous impact on your
likelihood of success and how you are viewed. Asking complete strangers for
help is often intimidating, especially the first time you do it. However, the
best way to do so is rarely taught explicitly --- at best, it's something one
picks up by example, from fellow lab members or the PI.  {\emph The goal of this
paper is to reveal the hidden curriculum -- what's expected of a software
consumer, how to ask for help, how to contribute productively to a software
community (regardless of whether you can write code).}

This article is written from a the perspective of a developer of academic
open-source scientific software, based on my personal experience as a developer
and user. I released my first piece of open source code, wham \cite{WHAM}, a
tool for analyzing umbrella sampling simulation,  during the summer of 2000.  My
group released LOOS\cite{Grossfield-2009, LOOS-JCC}, a suite of tools for
developing new tools to analyze molecular dynamics simulations, in 2008. Over
the years, I've interacted with hundreds of people who've tried to use one
package or the other. Most of these interactions have been positive, some
overwhelmingly so. Others were not. This article is an attempt to distill my
experiences into a concise set of recommendations.

Much of what I suggest is universal, but some points --- like the
reminder that the people providing the support are likely volunteers --- are
not.  Regardless, we view the interaction between developers and users as an
informal social contract, where each has obligations and expectations for the
other. Obviously, every software community is unique, and many have specific
conventions for interaction, but we hope that the recommendations we make are at
the very least a good starting point for new users of scientific software.

@@

Finally, you can consider diving into the code. This decision depends very much
on the complexity of the package, your skill level, and how urgently you need
the solution. More often than not, comprehending the whole software package will
be too time consuming, but there are tricks and shortcuts one can use to at
least figure out where things might be going wrong. Figuring out where the error
is occurring can sometimes help you figure out why it happened, so examine the
stack trace if there is one. Alternatively, you can search the code base for the
error message, to see what drove the error; it's not uncommon for the comments
in the code to be clearer and more indicative of what could go wrong than the
error message itself. To be clear: as an end user, \emph{you are not obligated
to check the code}, but it can sometimes be a good shortcut to help you either
solve your problem on your own or give the developers what they need to solve
it. Moreover, if you do solve the problem, it's a good idea to share your
solution with the developers; most will be grateful to you for your help.

\subsection{Ask for help in the right place}

\subsection{Write a good bug report}

\subsection{Contribute to the community}

\subsection{Treat people with respect}

This should go without saying, but if you're asking for help it behooves you to
do so respectfully. We're not saying that you need to be obsequious. Rather,
approach every interaction with respect: respect for people's time, respect for
their ideas, respect for their identity. If you're asking for help, do so with
the perspective that the developer is volunteering their time to help you, with
minimal payoff to them.  If you're reporting a bug, you're almost certainly
frustrated; try not to take that frustration out on the developers. If you're
offering a new feature, keep in mind that it may not align with the developers'
vision for where the software is going.

If you're interacting with fellow users in a support forum, don't mock
questioners who are less experienced than you --- a question may seem foolish to
you, but even an uninformed or ill-directed question is an opportunity for
teaching; one of our goals for this document is that it become a resource to
educate new users of expected norms without consuming developers' time and
mental bandwidth.

@@

Treating everyone with courtesy and respect, regardless of their identity, is
essential if you want the software tool to have a vibrant community surrounding
it. Ensuring people aren't excluded is the fair and just way to proceed ---
making sure that racist, sexist, anti-LGBTQ, etc language and behavior is
unwelcome and unacceptable is simply the moral thing to do (it's also pragmatic
--- can you really afford to lose talented developers and users because they
don't feel welcome?) This doesn't just mean use of epithets in messages (though
that's clearly unacceptable). Keep an eye on how things are named, on jokes
present in the code or messages, on whose opinions are listened to and who is
ignored.  Above all, if someone tells you a particular behavior or situation is
harmful, \emph{believe them}. Something that feels minor (or is just  a joke) to
you may hit others very differently, and part of respect is understanding and
accounting for that.

@@

\begin{checklist}{Good community member}
\begin{itemize}
\item Tries to solve problem themselves first
\item Asks for help in the right place
\item Writes informative bug reports
\item Cites and acknowledges software appropriately
\item Contributes to the community
\item Treats fellow members and developers with courtesy and respect
\end{itemize}
\end{checklist}
